### 2020-04-29 Rapport/chapters/Description_BPP.tex

\documentclass[class=article, crop=false]{standalone}
\usepackage[subpreambles=true]{standalone}
\usepackage{import}
\begin{document}
    Lorem ipsum....
    \section{Formulation Mathématique}
\end{document}